VOI-aware Monte-Carlo Sampling in Trees

### 2011-12-12 paper/vago/paper.tex
in progress

Bounds (\ref{eqn:bound-perf-hoeffding}, \ref{eqn:bound-blnk-hoeffding}) have the form
\begin{equation}
  \Lambda_i \le \hat\Lambda_i=\left\{
  \begin{array}{l l}
    (\overline X_\beta-B)\exp\left(- \gamma(\overline X_i - \overline X_\beta)^2 n_i \right) & \mbox{{\rm if} $i=\alpha$} \\
    (A-\overline  X_\alpha)\exp\left(- \gamma(\overline X_\alpha - \overline X_i)^2 n_i \right)  &  \mbox{\rm otherwise}
  \end{array} \right.
\label{eqn:bound-generic-hoeffding}
\end{equation}
where $A$ is the upper bound on the posterior sample mean $\overline
X_i'$ if $i\ne \alpha$, $B$ is the lower bound on $\overline X_i'$ if
$i=\alpha$. For $i=\alpha$, the bound can be improved by selecting an
intermediate point $B < y < \overline X_\beta$ and computing the bound
as the sum of two parts: 
\begin{itemize}
\item $\overline X_\beta-y$ multiplied by the probability that
  $\overline X_i' \le \overline X_\beta$;
\item $\overline X_\beta-B$ multiplied by the probability that $\overline X_i'\le
  y$.
\end{itemize}.
\begin{equation}
\Lambda_{i|i=\alpha} \le \hat \Lambda_{i|i=\alpha} =
 (\overline X_\beta-y)\exp\left(-\gamma (\overline X_i-\overline X_\beta)^2n_i\right)
 +(\overline X_\beta-B)\exp\left(-\gamma (\overline X_i-y)^2n_i\right)
\label{eqn:bound-twopart-hoeffding-alpha}
\end{equation}
If $y$ that minimizes $\hat\Lambda_{i|i=\alpha}$ exists, the minimum is achieved when
 $\frac {d\hat\Lambda_{i|i=\alpha}} {dy}=0$, that
is, when $y$ is the root of the following equation:
\begin{equation}
  2\gamma (\overline X_\beta-B)(\overline X_i-y)n_i = \exp\left(-\gamma \left(\frac {\overline X_i-\overline X_\beta}
    {\overline X_i-y}\right)^2n_i\right)
\end{equation}
The derivation for the case $i\ne \alpha$ is obtained by substitution $1-A, 1-\overline X_i, 1-\overline X_\alpha, 1-y$ instead of
$B, \overline X_i, \overline X_\beta, y$ into (\ref{eqn:bound-twopart-hoeffding-alpha}):
\begin{equation}
\Lambda_{i|i\ne\alpha} \le \hat\Lambda_{i|i\ne\alpha} = (y-\overline X_\alpha)\exp\left(-\gamma(\overline X_\alpha-\overline
  X_i)^2 n_i\right)+(A-\overline X_\alpha)\exp\left(-\gamma (y-\overline X_i)^2n_i\right)
\label{eqn:bound-twopart-hoeffding-rest}
\end{equation}
A closed-form solution for $y$ cannot be obtained, but given
$A, B, \overline X_\alpha, \overline X_\beta, n$, the value of $y$ can be efficiently
computed.

>>>>

Bounds (\ref{eqn:bound-perf-hoeffding}, \ref{eqn:bound-blnk-hoeffding}) have the form
\begin{equation}
  \Lambda_i \le \hat\Lambda_i=\left\{
  \begin{array}{l l}
    (\overline X_\beta-B)\exp\left(- \gamma(\overline X_i - \overline X_\beta)^2 n_i \right) & \mbox{{\rm if} $i=\alpha$} \\
    (A-\overline  X_\alpha)\exp\left(- \gamma(\overline X_\alpha - \overline X_i)^2 n_i \right)  &  \mbox{\rm otherwise}
  \end{array} \right.
\label{eqn:bound-generic-hoeffding}
\end{equation}
where $A$ is the upper bound on the posterior sample mean $\overline
X_i'$ if $i\ne \alpha$, $B$ is the lower bound on $\overline X_i'$ if
$i=\alpha$. For $i=\alpha$, the bound can be improved by selecting an
intermediate point $y$, $B < y < \overline X_\beta$, and computing the bound
as the sum of two parts: 
\begin{itemize}
\item $\overline X_\beta-y$ multiplied by the probability that
  $\overline X_i' \le \overline X_\beta$;
\item $\overline X_\beta-B$ multiplied by the probability that $\overline X_i'\le
  y$.
\end{itemize}.
\begin{equation}
\Lambda_{i|i=\alpha} \le \hat \Lambda_{i|i=\alpha} =
 (\overline X_\beta-y)\exp\left(-\gamma (\overline X_i-\overline X_\beta)^2n_i\right)
 +(\overline X_\beta-B)\exp\left(-\gamma (\overline X_i-y)^2n_i\right)
\label{eqn:bound-twopart-hoeffding-alpha}
\end{equation}
If $y$ that minimizes $\hat\Lambda_{i|i=\alpha}$ exists, the minimum is achieved when
 $\frac {d\hat\Lambda_{i|i=\alpha}} {dy}=0$, that
is, when $y$ is the root of the following equation:
\begin{equation}
  2\gamma (\overline X_\beta-B)(\overline X_i-y)n_i = \exp\left(-\gamma \left(\frac {\overline X_i-\overline X_\beta}
    {\overline X_i-y}\right)^2n_i\right)
\end{equation}
The derivation for the case $i\ne \alpha$ is obtained by substitution $1-A, 1-\overline X_i, 1-\overline X_\alpha, 1-y$ instead of
$B, \overline X_i, \overline X_\beta, y$ into (\ref{eqn:bound-twopart-hoeffding-alpha}):
\begin{equation}
\Lambda_{i|i\ne\alpha} \le \hat\Lambda_{i|i\ne\alpha} = (y-\overline X_\alpha)\exp\left(-\gamma(\overline X_\alpha-\overline
  X_i)^2 n_i\right)+(A-\overline X_\alpha)\exp\left(-\gamma (y-\overline X_i)^2n_i\right)
\label{eqn:bound-twopart-hoeffding-rest}
\end{equation}
A closed-form solution for $y$ cannot be obtained, but given
$A, B, \overline X_\alpha, \overline X_\beta, n$, the value of $y$ can be efficiently
computed.
